---
title: "Data Analysis"
output: html_document
---

```{r}
library(tidyverse)
library(survey)
library(viridis)
library(table1)
library(kableExtra)

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d(option = "viridis")
scale_fill_discrete = scale_fill_viridis_d(option = "viridis")
```



```{r, functions}
drug_use_tidy <- function(input_list){
  output <- case_when(input_list == 1 ~ 0,
                      input_list == 2 ~ 1.5,
                      input_list == 3 ~ 6,
                      input_list == 4 ~ 14.5,
                      input_list == 5 ~ 29.5,
                      input_list > 5 ~ 40)
  return(output)
}
drug_use_frequency <- function(input_list){
  output <- case_when(input_list == 0 ~ "No Drug Use",
                      input_list > 0 & input_list < 40 ~ "Light Dose",
                      input_list >= 40 ~ "Heavy Dose")
  factor(output, levels = c("No Drug Use", "Light Dose", "Heavy Dose"))
}

age_tidy <- function(input_list) {
  output <- case_when(input_list == 1 ~ NA_real_,
                      input_list == 2 ~ 8,
                      input_list == 3 ~ 9.5,
                      input_list == 4 ~ 11.5,
                      input_list == 5 ~ 13.5,
                      input_list == 6 ~ 15.5,
                      input_list == 7 ~ 17.5)
}

firstage_tidy <- function(input_list){
  output <- case_when(input_list == 1 ~ "never",
                      input_list == 2 ~ "8 years old or younger",
                      input_list == 3 ~ "9 or 10 years old",
                      input_list == 4 ~ "11 or 12 years old",
                      input_list == 5 ~ "13 or 14 years old",
                      input_list == 6 ~ "15 or 16 years old",
                      input_list == 7 ~ "17 years old or older")
  return(output)
}


association_analysis <- function(input_df, input_name, type){
  if (type == "crude") {
    totaln <- input_df %>% nrow()
    output <- 
      input_df %>% 
      group_by(drug_use_freq, across(input_name)) %>%
      count() %>%
      pivot_wider(values_from = n, names_from = starts_with(input_name)) %>%
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      chisq.test() %>% 
      broom::tidy() %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter),
             stratum = "Overall") %>% 
      select(stratum, statistic, p.value, df = parameter, cramer_v_effect_size)
  } else {
    output <- 
      input_df %>% 
      group_by(across(all_of(type)), drug_use_freq, across(all_of(input_name))) %>%
      count() %>% 
      pivot_wider(values_from = n, names_from = starts_with(input_name), names_prefix = input_name) %>% 
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      nest(data = starts_with(input_name)) %>% 
      mutate(
      chisq_tests = map(.x = data, ~chisq.test(x = .x)),
      chisq_result = map(chisq_tests, broom::tidy),
      totaln = map(data, colSums),
      totaln = map_dbl(.x = totaln, ~reduce(.x, sum))) %>% 
      unnest(chisq_result) %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter)) %>% 
      select(stratum = starts_with(type), 
             statistic, p.value, 
             df = parameter, 
             cramer_v_effect_size)
  }
  return(output)
}

test_of_association <- function(input_df, input_name, table_name){
  output <- tibble(
    type = c("crude", "grade", "sex", "race4")
  )
  output <- output %>% 
    mutate(results = map(.x = type, ~association_analysis(input_df, input_name, type = .x)),
           type = ifelse(type == "race4", "race", type),
           type = factor(type, levels = c("crude","grade", "sex", "race"))) %>% 
    unnest(results)
  kbl(output[, -1], caption = table_name) %>%
    kable_paper("hover", full_width = F) %>% 
    pack_rows(index = table(output$type))
}

```

```{r import data}
yrbss_state_a <- readRDS("data/2019 states a-m.rds")
yrbss_state_z <- readRDS("data/2019 states n-z.rds")
yrbss_state = rbind(yrbss_state_a,yrbss_state_z)
yrbss_state = as_tibble(yrbss_state)
```


```{r tidy data}
yrbss_state_tidy <-
  yrbss_state %>% 
  drop_na(q31, q40, q45, q46, q50, q51, q52, q53, q54, q55, q56, q9, q10, q14, q17, q26, q39,q59, q79, q80, q88, q33, q42, q89, q8, q11, grade, race4, sex) %>% 
  mutate(
    grade = fct_recode(factor(grade, ordered = TRUE), 
                       "9th Grade" = "1",
                       "10th Grade" = "2",
                       "11th Grade" = "3",
                       "12th Grade" = "4"),
    sex = fct_recode(factor(sex, ordered = TRUE), 
                       "female" = "1",
                       "male" = "2"),
    race4 = fct_recode(factor(race4), 
                       "White" = "1",
                       "Black or African American" = "2",
                       "Hispanic/Latino" = "3",
                       "All other Races" = "4"),
    age_initial_alcohol = age_tidy(q40),
    age_initial_smoking = age_tidy(q31),
    age_marijuana = age_tidy(q46),
    marijuana = drug_use_tidy(q45),
    cocaine = drug_use_tidy(q50),
    inhale = drug_use_tidy(q51),
    heroin = drug_use_tidy(q52),
    methamphetamine = drug_use_tidy(q53),
    ecstasy = drug_use_tidy(q54),
    steroid = drug_use_tidy(q55),
    illegal_injection = case_when(q56 == 1 ~ 0,
                                  q56 == 2 ~ 20,
                                  q56 == 3 ~ 40),
    drug_use_sum = rowSums(across(marijuana:illegal_injection)),
    other_drug_sum = rowSums(across(cocaine:illegal_injection)),
    drug_use_freq = drug_use_frequency(drug_use_sum),
    drunk_driving = ifelse(q9 > 1 | q10 > 2, "Yes", "No"),
    drunk_driving = factor(drunk_driving, level = c("Yes", "No")),
    text_driving = ifelse(q11 == 2, "No", "Yes"),
    text_driving = factor(text_driving, level = c("Yes", "No")),
    carrying_weapon = ifelse(q14 == 1, "No", "Yes"),
    carrying_weapon = factor(carrying_weapon, level = c("Yes", "No")),
    suicide_attempt = ifelse(q26 == 1, "Yes", "No"),
    suicide_attempt = factor(suicide_attempt, level = c("Yes", "No")),
    quit_smoke = case_when(q39 == 1 ~ "Never Smoke",
                           q39 == 2 ~ "Yes",
                           q39 == 3 ~ "No"),
    quit_smoke = factor(quit_smoke, level = c("Never Smoke", "Yes", "No")),
    physical_fight = ifelse(q17 == 1, "No", "Yes"),
    physical_fight = factor(physical_fight, level = c("Yes", "No")),
    early_sex = ifelse(q59 > 1 & q59 < 5 , "Yes", "No"),
    early_sex = factor(early_sex, level = c("Yes", "No")),
    q79 = as.numeric(q79),
    q80 = as.numeric(q80),
    tv_use = case_when(q79 == 1 ~ 0, 
                       q79 == 2 ~ 0.5,
                       q79 > 2 ~ q79 - 2),
    computer_use = q80 - 2,
    computer_use = case_when(q80 == 1 ~ 0, 
                       q80 == 2 ~ 0.5,
                       q80 > 2 ~ q80 - 2),
    screening_use = tv_use + computer_use,
    sleeping_time = as.numeric(q88) + 3,
    smoking_status = case_when(q33 == 1 ~ "Never Smoker",
                               q33 >= 2 & q33 <= 5 ~ "Light Smoker",
                               q33 >=6 ~ "Heavy Smoker"),
    smoking_status = factor(smoking_status, level = c("Never Smoker", "Light Smoker", "Heavy Smoker")),
    binge_drinking = case_when(q42 == 1 ~ "No Binge Drinking",
                               q42 >= 2 & q42 <= 4 ~ "Light Binge Drinking",
                               q42 >=5 ~ "Heavy Binge Drinking"),
    binge_drinking = factor(binge_drinking, level = c("No Binge Drinking", "Light Binge Drinking", "Heavy Binge Drinking")),
    seat_belt = case_when(q8 <= 2 ~ "Never or Rarely",
                          q8 == 3 ~ "Sometimes",
                          q8 >= 4 ~ "Most of the time or Always"),
    seat_belt = factor(seat_belt, level = c("Never or Rarely", "Sometimes", "Most of the time or Always")),
    grades_school = case_when(q89 == 1 ~ "Mostly A's",
                       q89 == 2 ~ "Mostly B's",
                       q89 == 3 ~ "Mostly C's",
                       q89 %in% c(4,5) ~ "Mostly Below C's",
                       q89 == 6 ~ "None of these grades",
                       q89 == 7 ~ "Not sure"),
    grades_school = factor(grades_school, level = c("Mostly A's", "Mostly B's", "Mostly C's", "Mostly Below C's", "None of these grades", "Not sure")),
    smoke_age = firstage_tidy(q31),
    alcohol_age = firstage_tidy(q40)
    ) %>% 
  select(drug_use_freq, grade, sex, bmi, race4, drunk_driving, text_driving, carrying_weapon, suicide_attempt, quit_smoke, physical_fight, early_sex, screening_use, sleeping_time, smoking_status, binge_drinking, seat_belt, grades_school, smoke_age, alcohol_age,)
```

This data analysis uses the resulted data set cleaned by the processing described in part 2 of [Data]("3,Data Description.html"). According to our motivation, we are interested the relationship between drug use and youth behaviors. In this analysis, we will explore the association between drug use status and each health behavior variables.

There are three kinds of health-related variables and we will use different methods to analyze: 
 - Bi-level categorical variables: drunk driving, carrying weapon, suicide attempt, physical fight, quit-smoking, early-age sexual intercourse, heavy screening usage, enough sleeping
 - Multivariate categorical variables: smoking status, binge drinking status, seat belt usage, GPA
 - Quantitative Variables: initial smoking age, initial drinking age, BMI
 
Here is a table summarized the overall distribution of important variables by drug use status.

## Three-line-table

```{r three-line}
table1(~ drunk_driving + carrying_weapon + suicide_attempt + quit_smoke + physical_fight + early_sex + screening_use + smoking_status + binge_drinking + seat_belt + grades_school| drug_use_freq, data = df, caption = "Table 1: Distribution of risky behaviors by drug use status")
```



## Hypothesis Testing with visualization of drug status comparison

### Chi-square Test

For bi-level and multi-level categorical variables, we test the independence of two categorical variables, so we should use the R X C chi-square test with hypotheses that 

$$
H_0:\text{There is no association between drug use and the health-related variable of interest}
H_1:\text{There is an association between drug use and the health-related variable of interest}
$$
Besides the chi-square test, we also use the Cramer's V Effect Size to measure the magnitude of the association. It is measure bt the manner:
$$
V = \sqrt{\frac{X^2}{n*df}}
$$
where $X^2$ is the chi-square statistic, n is the sample size and $df$ is the degrees of freedom.\

Usually, we use the table below to assess the association through the Cramer's V effect size

```{r}
tibble(
  df = 1:5,
  small_effect = c(0.10, 0.07, 0.06, 0.05, 0.04),
  medium_effect = c(0.3, 0.21, 0.17, 0.15, 0.13),
  large_effect = c(0.5, 0.35, 0.29, 0.25, 0.22)
) %>% 
  knitr::kable()
```



### Drunk driving

```{r}
test_of_association(yrbss_state_tidy, "drunk_driving", "Table 2: Drug Use Status vs Drunk Driving")
```


```{r}
yrbss_state_tidy_drunk_driving %>% 
  group_by(drug_use_freq) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_drunk_driving %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_drunk_driving %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_drunk_driving %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```


### cell phone use while driving
```{r}
yrbss_state_tidy_text_driving <- yrbss_state_tidy %>%
  drop_na(q11) %>% 
  mutate(
    text_driving = ifelse(q11 == 2, "No", "Yes")
  )
```


```{r}
test_of_association(yrbss_state_tidy_text_driving, "text_driving")

yrbss_state_tidy_text_driving %>% 
  group_by(drug_use_freq) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_text_driving %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_text_driving %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_text_driving %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```



### Carrying weapon

```{r}
yrbss_state_tidy_carrying_weapon <- yrbss_state_tidy %>%
  drop_na(q14) %>% 
  mutate(
    carrying_weapon = ifelse(q14 == 1, "No", "Yes")
  )
```


```{r}
test_of_association(yrbss_state_tidy_carrying_weapon, "carrying_weapon")

yrbss_state_tidy_carrying_weapon %>% 
  group_by(drug_use_freq) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_carrying_weapon %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_carrying_weapon %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_carrying_weapon %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```

```{r}
yrbss_state_tidy %>% 
  filter(year == 2019) %>% 
  drop_na(q14) %>% 
  mutate(
    carrying_weapon = ifelse(q14 == 1, "No", "Yes"),
    drug_usage = case_when(other_drug_sum == 0 & marijuana > 0 ~ "marijuana",
                                other_drug_sum > 0 & marijuana == 0 ~ "other drug",
                                other_drug_sum > 0 & marijuana > 0 ~ "Both drug",
                           other_drug_sum == 0 & marijuana == 0 ~ "No drug")
  ) %>% 
  filter(drug_usage %in% c("marijuana", "other drug")) %>% 
  group_by(drug_usage, carrying_weapon) %>% 
  count() %>% 
  pivot_wider(values_from = n, names_from = carrying_weapon, names_prefix = "carrying_weapon_") %>% 
  ungroup() %>%
  select(-drug_usage) %>% 
  chisq.test()

yrbss_state_tidy %>% 
  drop_na(q14) %>% 
  mutate(
    carrying_weapon = ifelse(q14 == 1, "No", "Yes"),
    drug_usage = case_when(other_drug_sum == 0 & marijuana > 0 ~ "marijuana",
                                other_drug_sum > 0 & marijuana == 0 ~ "other drug",
                                other_drug_sum > 0 & marijuana > 0 ~ "Both drug",
                           other_drug_sum == 0 & marijuana == 0 ~ "No drug")
  ) %>% 
  filter(drug_usage %in% c("marijuana", "other drug")) %>% 
  group_by(drug_usage) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_usage, y = prop_carrying_weapon)) +
  geom_bar(stat = "identity", position = "dodge")
```



### Suicide Attempt

```{r}
yrbss_state_tidy_suicide_attempt <- 
  yrbss_state_tidy %>%
  drop_na(q26) %>% 
  mutate(
    suicide_attempt = ifelse(q26 == 1, "Yes", "No")
  )
```


```{r}
test_of_association(yrbss_state_tidy_suicide_attempt, "suicide_attempt")

yrbss_state_tidy_suicide_attempt %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_suicide_attempt %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_suicide_attempt %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_suicide_attempt %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```


```{r}
yrbss_state_tidy %>% 
  filter(year == 2019) %>% 
  drop_na(q26) %>% 
  mutate(
    suicide_attempt = ifelse(q26 == 1, "Yes", "No"),
    drug_usage = case_when(other_drug_sum == 0 & marijuana > 0 ~ "marijuana",
                                other_drug_sum > 0 & marijuana == 0 ~ "other drug",
                                other_drug_sum > 0 & marijuana > 0 ~ "Both drug",
                           other_drug_sum == 0 & marijuana == 0 ~ "No drug")
  ) %>% 
  filter(drug_usage %in% c("marijuana", "other drug")) %>% 
  group_by(drug_usage, suicide_attempt) %>% 
  count() %>% 
  pivot_wider(values_from = n, names_from = suicide_attempt, names_prefix = "suicide_attempt_") %>% 
  ungroup() %>%
  select(-drug_usage) %>% 
  chisq.test()

yrbss_state_tidy %>% 
  filter(year == 2019) %>% 
  drop_na(q26) %>% 
  mutate(
    suicide_attempt = ifelse(q26 == 1, "Yes", "No"),
    drug_usage = case_when(other_drug_sum == 0 & marijuana > 0 ~ "marijuana",
                                other_drug_sum > 0 & marijuana == 0 ~ "other drug",
                                other_drug_sum > 0 & marijuana > 0 ~ "Both drug",
                           other_drug_sum == 0 & marijuana == 0 ~ "No drug")
  ) %>% 
  filter(drug_usage %in% c("marijuana", "other drug")) %>% 
  group_by(drug_usage) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_usage, y = prop_suicide_attempt)) +
  geom_bar(stat = "identity", position = "dodge")
```

### Quit smoking

```{r}
yrbss_state_tidy_quit_smoke <- 
  yrbss_state_tidy %>%
  drop_na(q39) %>% 
  filter(q39 != 1) %>% 
  mutate(
    quit_smoke = case_when(q39 == 2 ~ "Yes",
                           q39 == 3 ~ "No")
  )
```



```{r}
test_of_association(yrbss_state_tidy_quit_smoke, "quit_smoke")

yrbss_state_tidy_quit_smoke %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_quit_smoke %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")

yrbss_state_tidy_quit_smoke %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")

race_analysis(yrbss_state_tidy_quit_smoke, "quit_smoke")

yrbss_state_tidy_quit_smoke %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```



### physical fight

```{r}
yrbss_state_tidy_physical_fight <- 
  yrbss_state_tidy %>%
  drop_na(q17) %>% 
  mutate(
    physical_fight = ifelse(q17 == 1, "No", "Yes")
  )
```



```{r}
test_of_association(yrbss_state_tidy_physical_fight, "physical_fight")

yrbss_state_tidy_physical_fight %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_physical_fight %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_physical_fight %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_physical_fight %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```


### Early-age Sexual Intercourse

```{r}
yrbss_state_tidy_early_sex <- 
  yrbss_state_tidy %>%
  drop_na(q59) %>% 
  mutate(
    early_sex = ifelse(q59 > 1 & q59 < 5 , "Yes", "No")
  ) 
```



```{r}
test_of_association(yrbss_state_tidy_early_sex, "early_sex")

yrbss_state_tidy_early_sex %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_early_sex %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_early_sex %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_early_sex %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")

```



### Screening Use

```{r}
yrbss_state_tidy_screening_use <- 
  yrbss_state_tidy %>%
  drop_na(q79, q80) %>% 
  mutate(
    q79 = as.numeric(q79),
    q80 = as.numeric(q80),
    tv_use = case_when(q79 == 1 ~ 0, 
                       q79 == 2 ~ 0.5,
                       q79 > 2 ~ q79 - 2),
    computer_use = q80 - 2,
    computer_use = case_when(q80 == 1 ~ 0, 
                       q80 == 2 ~ 0.5,
                       q80 > 2 ~ q80 - 2),
    screening_use = tv_use + computer_use
  )
yrbss_state_tidy_screening_use %>% 
  group_by(drug_use_freq, screening_use) %>% 
  count() %>% 
  group_by(drug_use_freq) %>% 
  mutate(
    prop = n/sum(n)
  ) %>% 
  ggplot(aes(x = screening_use, y = prop)) +
  geom_bar(stat = "identity") +
  geom_smooth(se = FALSE) +
  facet_grid(.~drug_use_freq)
```


```{r}
yrbss_state_tidy_heavy_screen_use <-
  yrbss_state_tidy_screening_use %>% 
  mutate(
    heavy_screen_use = ifelse(screening_use == 5 | screening_use > 5, "Yes", "No")
  )


test_of_association(yrbss_state_tidy_heavy_screen_use, "heavy_screen_use")

yrbss_state_tidy_heavy_screen_use %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_heavy_screen_use %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_heavy_screen_use %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_heavy_screen_use %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```


### Sleeping time

```{r}
yrbss_state_tidy_sleeping_time <- 
  yrbss_state_tidy %>%
  drop_na(q88) %>% 
  mutate(
    sleeping_time = as.numeric(q88) + 3
  )

yrbss_state_tidy_sleeping_time %>% 
  group_by(drug_use_freq, sleeping_time) %>% 
  count() %>% 
  group_by(drug_use_freq) %>% 
  mutate(
    prop = n/sum(n)
  ) %>% 
  ggplot(aes(x = sleeping_time, y = prop)) +
  geom_bar(stat = "identity") +
  geom_smooth(se = FALSE) +
  facet_grid(.~drug_use_freq)
```


```{r}
yrbss_state_tidy_enough_sleep <-
  yrbss_state_tidy_sleeping_time %>% 
  mutate(
    enough_sleep = ifelse(sleeping_time == 8 | sleeping_time > 8, "Yes", "No")
  )

test_of_association(yrbss_state_tidy_enough_sleep, "enough_sleep")

yrbss_state_tidy_enough_sleep %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_enough_sleep %>% 
  mutate(grade = ifelse(is.na(grade), "Missing", as.character(grade))) %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_enough_sleep %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge")


yrbss_state_tidy_enough_sleep %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge")
```

### Smoking Status

Based on question 33 in the 2019 YOUTH RISK BEHAVIOR SURVEY, the researchers care about whether students smoked more than 10 cigarettes per day on the days they smoked during the 30 days before the survey. Therefore, we define students who did not smoke as "Never Smoker", 1 to 10 cigarettes per day as "Light Smoker", greater than 10 cigarettes per day as "Heavy Smokers".

```{r}
smoking_status_crude = 
  association_analysis(yrbss_state_tidy, "smoking_status", "crude")

# chi-square of grade
smoking_status_grade =
  yrbss_state_tidy %>% 
      group_by(across(all_of("grade")), drug_use_freq, across(all_of("smoking_status"))) %>%
      count() %>% 
      pivot_wider(values_from = n, names_from = starts_with("smoking_status"), names_prefix = "smoking_status") %>% 
  mutate(`smoking_statusHeavy Smoker` = ifelse(is.na(`smoking_statusHeavy Smoker`), 0 , `smoking_statusHeavy Smoker`)) %>% 
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      nest(data = starts_with("smoking_status")) %>% 
      mutate(
      chisq_tests = map(.x = data, ~chisq.test(x = .x)),
      chisq_result = map(chisq_tests, broom::tidy),
      totaln = map(data, colSums),
      totaln = map_dbl(.x = totaln, ~reduce(.x, sum))) %>% 
      unnest(chisq_result) %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter)) %>% 
      select(stratum = starts_with("grade"), 
             statistic, p.value, 
             df = parameter, 
             cramer_v_effect_size)

# chi-square of gender
smoking_status_gender = 
  association_analysis(yrbss_state_tidy, "smoking_status", "sex")

# chi-square of race
smoking_status_race = 
  yrbss_state_tidy %>% 
      group_by(across(all_of("race4")), drug_use_freq, across(all_of("smoking_status"))) %>%
      count() %>% 
      pivot_wider(values_from = n, names_from = starts_with("smoking_status"), names_prefix = "smoking_status") %>% 
  mutate(`smoking_statusHeavy Smoker` = ifelse(is.na(`smoking_statusHeavy Smoker`), 0 , `smoking_statusHeavy Smoker`)) %>% 
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      nest(data = starts_with("smoking_status")) %>% 
      mutate(
      chisq_tests = map(.x = data, ~chisq.test(x = .x)),
      chisq_result = map(chisq_tests, broom::tidy),
      totaln = map(data, colSums),
      totaln = map_dbl(.x = totaln, ~reduce(.x, sum))) %>% 
      unnest(chisq_result) %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter)) %>% 
      select(stratum = starts_with("race4"), 
             statistic, p.value, 
             df = parameter, 
             cramer_v_effect_size)
```

```{r}
yrbss_state_tidy %>% 
  group_by(drug_use_freq, smoking_status) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
 # scale_color_manual(values=c('#56B4E9', '#E69F00', '#999999'))
  labs(title = "Proportion of Smoking Status Among Drug Use Frequency",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Stratified Analysis over Grade
yrbss_state_tidy %>% 
  group_by(drug_use_freq,grade, smoking_status) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ grade) +
  labs(title = "Stratified Analysis of Smoking Stauts over Grade",
       x = "Drug Use Frequency",
       y = "Proportion") 
```

```{r}
# Over Gender
yrbss_state_tidy %>% 
  group_by(drug_use_freq,sex, smoking_status,) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ sex)
  labs(title = "Stratified Analysis of Smoking Stauts over Gender",
       x = "Drug Use Frequency",
       y = "Proportion") 
```




```{r}
# Over race
yrbss_state_tidy %>% 
  group_by(drug_use_freq, race4, smoking_status) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ race4) + 
  labs(title = "Stratified Analysis of Smoking Stauts over Race",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


### Binge Drinking 

Based on question 42 in the 2019 YOUTH RISK BEHAVIOR SURVEY, we define students who did not have binge drinking as "No Binge Drinking", 1 to 5 days of binge drinking in the past 30 days as "Light Binge Drinking", greater than 5days as "Heavy Binge Drinking".

```{r}
test_of_association(yrbss_state_tidy, "binge_drinking")
```

```{r}
yrbss_state_tidy %>% 
  group_by(drug_use_freq, binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Proportion of Binge Drinking Among Drug Use Frequency",
       x = "Drug Use Frequency",
       y = "Proportion")
```


```{r}
# Stratified Analysis over Grade
yrbss_state_tidy %>% 
  group_by(drug_use_freq,grade, binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ grade) + 
  labs(title = "Stratified Analysis of Binge Drinking over Grade",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


```{r}
# Over Gender
yrbss_state_tidy %>% 
  group_by(drug_use_freq,sex,binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ sex) +
  labs(title = "Stratified Analysis of Binge Drinking over Gender",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


```{r}
# Over race
yrbss_state_tidy %>% 
  group_by(drug_use_freq, race4, binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ race4) + 
  labs(title = "Stratified Analysis of Binge Drinking over Race",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


### Seat belt use 
Resource: question 8

```{r}
test_of_association(yrbss_state_tidy, "seat_belt")
```

```{r}
yrbss_state_tidy %>% 
  group_by(drug_use_freq, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Proportion of Seat Belt Use Among Drug Use Frequency",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Stratified Analysis over Grade
yrbss_state_tidy%>% 
  group_by(drug_use_freq,grade, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ grade) +
  labs(title = "Stratified Analysis of Seat Belt Use over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Over Gender
yrbss_state_tidy %>% 
  group_by(drug_use_freq,sex, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ sex) +
  labs(title = "Stratified Analysis of Seat Belt Use over Gender",
       x = "Drug Use Frequency",
       y = "Proportion") 
```



```{r}
# Over race
yrbss_state_tidy %>% 
  group_by(drug_use_freq, race4, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ race4) + 
  labs(title = "Stratified Analysis of Seat Belt Use over Race",
       x = "Drug Use Frequency",
       y = "Proportion")
```

### Grades in school 
Resource: question 89
exclude option "None of these grades" and "Not sure"

```{r}
yrbss_state_tidy_grades_school <- 
  yrbss_state_tidy %>% 
  filter(grades_school %in% c("Mostly A's", "Mostly B's", "Mostly C's","Mostly Below C's"))
```

```{r}
test_of_association(yrbss_state_tidy_grades_school, "grades_school")
```

```{r}
yrbss_state_tidy_grades_school %>% 
  group_by(drug_use_freq, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Proportion of Grades in School Among Drug Use Frequency",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Stratified Analysis over Grade
yrbss_state_tidy_grades_school %>% 
  group_by(drug_use_freq,grade, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(.~ grade) +
  labs(title = "Stratified Analysis of Grades in School over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Over Gender
yrbss_state_tidy_grades_school %>% 
  group_by(drug_use_freq,sex, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ sex) +
  labs(title = "Stratified Analysis of Grades in School over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Over race
yrbss_state_tidy_grades_school %>% 
  group_by(drug_use_freq, race4, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid( . ~ race4) +
  labs(title = "Stratified Analysis of Grades in School over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```




## Part 3
##### 2，额外图  
1，吸毒各个frequency段的，吸烟初始年龄分布。  
2, .....................，喝酒初始年龄分布。  
3, ....................., bmi box plot 分男女 or 金字塔图

## Q1
```{r}
#tidy q31 smoking and q40 alcohol start age
firstage_tidy <- function(input_list){
  output <- case_when(input_list == 1 ~ "never",
                      input_list == 2 ~ "8 years old or younger",
                      input_list == 3 ~ "9 or 10 years old",
                      input_list == 4 ~ "11 or 12 years old",
                      input_list == 5 ~ "13 or 14 years old",
                      input_list == 6 ~ "15 or 16 years old",
                      input_list == 7 ~ "17 years old or older")
  return(output)
}

#function for plotting first smoke_age
drug1_smoke=
  yrbss_state_tidy %>%
  mutate(
    smoke_age = firstage_tidy(q31),
  ) %>%
  drop_na(q31)


drug1_smoke %>%
  group_by(drug_use_freq, smoke_age) %>%
  filter(smoke_age != "never") %>%
  summarize(n = n()) %>% 
  mutate(percentage = n/sum(n),
         percent = scales::percent(percentage),
         smoke_age = factor(smoke_age, 
                            levels = c("never",
                                       "8 years old or younger",
                                       "9 or 10 years old",
                                       "11 or 12 years old",
                                       "13 or 14 years old",
                                       "15 or 16 years old",
                                       "17 years old or older"))) %>% 
  ggplot(
       aes(x = drug_use_freq, y = percentage,
           fill = smoke_age)
       ) +
  geom_bar(stat = "identity",                                     
           position = "dodge") +
  scale_y_continuous(breaks = seq(0, 2, .1), 
                     label = scales::percent) +
  geom_text(aes(label = percent),                                     
            size = 3,
            position = position_dodge(0.9)) +
  scale_fill_brewer(palette = "Set2")+
  labs(title="smoke start age")


#boxplot
ggplot(drug1_smoke, 
       aes(x = factor(drug_use_freq), 
           y = age_initial_smoking,fill=factor(drug_use_freq))) +     
  geom_boxplot() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1),
    legend.position = "none") +
  labs(title="smoke start age")
```

## Q2-alcohol
```{r}
#function for plotting first alcohol_age
drug1_alcohol=
  yrbss_state_tidy %>%
  mutate(
    alcohol_age = firstage_tidy(q40)
  ) %>%
  drop_na(q40) %>%
  drop_na(drug_use_freq)


drug1_alcohol %>%
  group_by(drug_use_freq, alcohol_age) %>%
    filter(alcohol_age != "never") %>%
  summarize(n = n()) %>% 
  mutate(percentage = n/sum(n),
         percent = scales::percent(percentage),
         alcohol_age = factor(alcohol_age, 
                            levels = c("never",
                                       "8 years old or younger",
                                       "9 or 10 years old",
                                       "11 or 12 years old",
                                       "13 or 14 years old",
                                       "15 or 16 years old",
                                       "17 years old or older"))) %>% 
  ggplot(
       aes(x = drug_use_freq, y = percentage,
           fill = alcohol_age)
       ) +
  geom_bar(stat = "identity",                                     
           position = "dodge") +
  scale_y_continuous(breaks = seq(0, 2, .1), 
                     label = scales::percent) +
  geom_text(aes(label = percent),                                     
            size = 3,
            position = position_dodge(0.9)) +
  scale_fill_brewer(palette = "Set2")+
  labs(title="alcohol start age")

#boxplot
ggplot(drug1_alcohol, 
       aes(x = factor(drug_use_freq), 
           y = age_initial_alcohol,fill=factor(drug_use_freq))) +        
  geom_boxplot() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1),
    legend.position = "none") +
  labs(title="alcohol start age")

```

## Q3-BMI
```{r}
BMI =
  yrbss_state_tidy %>%
  drop_na(drug_use_freq) %>%
  drop_na(sex)

BMI %>%
  ggplot(aes(x = factor(drug_use_freq), 
           y = bmi,fill=factor(drug_use_freq))) +
  geom_boxplot() +
  facet_grid(. ~ sex) +  
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1),
    legend.position = "none")+
  labs(title = "Salary distribution by rank")

#heavy tail - more dose have higher BMI 


#pyramid plot - looks weird
ggplot(BMI, aes(x = drug_use_freq, y = ifelse(sex == "2", -bmi, bmi), fill =sex)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = abs, limits = c(-200,200)) +
  coord_flip()

```



## Logistic regression
