---
title: "Data Analysis"
output: html_document
---

```{r}
library(tidyverse)
library(survey)
library(viridis)
library(table1)
library(kableExtra)

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  echo = FALSE
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d(option = "viridis")
scale_fill_discrete = scale_fill_viridis_d(option = "viridis")
```



```{r, functions}
drug_use_tidy <- function(input_list){
  output <- case_when(input_list == 1 ~ 0,
                      input_list == 2 ~ 1.5,
                      input_list == 3 ~ 6,
                      input_list == 4 ~ 14.5,
                      input_list == 5 ~ 29.5,
                      input_list > 5 ~ 40)
  return(output)
}
drug_use_frequency <- function(input_list){
  output <- case_when(input_list == 0 ~ "No Drug Use",
                      input_list > 0 & input_list < 40 ~ "Light Dose",
                      input_list >= 40 ~ "Heavy Dose")
  factor(output, levels = c("No Drug Use", "Light Dose", "Heavy Dose"))
}

age_tidy <- function(input_list) {
  output <- case_when(input_list == 1 ~ NA_real_,
                      input_list == 2 ~ 8,
                      input_list == 3 ~ 9.5,
                      input_list == 4 ~ 11.5,
                      input_list == 5 ~ 13.5,
                      input_list == 6 ~ 15.5,
                      input_list == 7 ~ 17.5)
}

firstage_tidy <- function(input_list){
  output <- case_when(input_list == 1 ~ "never",
                      input_list == 2 ~ "8 years old or younger",
                      input_list == 3 ~ "9 or 10 years old",
                      input_list == 4 ~ "11 or 12 years old",
                      input_list == 5 ~ "13 or 14 years old",
                      input_list == 6 ~ "15 or 16 years old",
                      input_list == 7 ~ "17 years old or older")
  return(output)
}


association_analysis <- function(input_df, input_name, type){
  if (type == "crude") {
    totaln <- input_df %>% nrow()
    output <- 
      input_df %>% 
      group_by(drug_use_freq, across(input_name)) %>%
      count() %>%
      pivot_wider(values_from = n, names_from = starts_with(input_name)) %>%
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      chisq.test() %>% 
      broom::tidy() %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter),
             stratum = "Overall") %>% 
      select(stratum, statistic, p.value, df = parameter, cramer_v_effect_size)
  } else {
    output <- 
      input_df %>% 
      group_by(across(all_of(type)), drug_use_freq, across(all_of(input_name))) %>%
      count() %>% 
      pivot_wider(values_from = n, names_from = starts_with(input_name), names_prefix = input_name) %>% 
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      nest(data = starts_with(input_name)) %>% 
      mutate(
      chisq_tests = map(.x = data, ~chisq.test(x = .x)),
      chisq_result = map(chisq_tests, broom::tidy),
      totaln = map(data, colSums),
      totaln = map_dbl(.x = totaln, ~reduce(.x, sum))) %>% 
      unnest(chisq_result) %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter)) %>% 
      select(stratum = starts_with(type), 
             statistic, p.value, 
             df = parameter, 
             cramer_v_effect_size)
  }
  return(output)
}

test_of_association <- function(input_df, input_name, table_name){
  output <- tibble(
    type = c("crude", "grade", "sex", "race4")
  )
  output <- output %>% 
    mutate(results = map(.x = type, ~association_analysis(input_df, input_name, type = .x)),
           type = ifelse(type == "race4", "race", type),
           type = factor(type, levels = c("crude","grade", "sex", "race"))) %>% 
    unnest(results)
  kbl(output[, -1], caption = table_name) %>%
    kable_paper("hover", full_width = F) %>% 
    pack_rows(index = table(output$type))
}

```

```{r import data}
yrbss_state_a <- readRDS("data/2019 states a-m.rds")
yrbss_state_z <- readRDS("data/2019 states n-z.rds")
yrbss_state = rbind(yrbss_state_a,yrbss_state_z)
yrbss_state = as_tibble(yrbss_state)
```


```{r tidy data}
yrbss_state_tidy_additional <-
  yrbss_state %>% 
  drop_na(q31, q40, q45, q46, q50, q51, q52, q53, q54, q55, q56, q9, q10, q14, q17, q26, q39,q59, q79, q80, q88, q33, q42, q89, q8, q11, grade, race4, sex) %>% 
  mutate(
    grade = fct_recode(factor(grade, ordered = TRUE), 
                       "9th Grade" = "1",
                       "10th Grade" = "2",
                       "11th Grade" = "3",
                       "12th Grade" = "4"),
    sex = fct_recode(factor(sex, ordered = TRUE), 
                       "female" = "1",
                       "male" = "2"),
    race4 = fct_recode(factor(race4), 
                       "White" = "1",
                       "Black or African American" = "2",
                       "Hispanic/Latino" = "3",
                       "All other Races" = "4"),
    age_initial_alcohol = age_tidy(q40),
    age_initial_smoking = age_tidy(q31),
    age_marijuana = age_tidy(q46),
    marijuana = drug_use_tidy(q45),
    cocaine = drug_use_tidy(q50),
    inhale = drug_use_tidy(q51),
    heroin = drug_use_tidy(q52),
    methamphetamine = drug_use_tidy(q53),
    ecstasy = drug_use_tidy(q54),
    steroid = drug_use_tidy(q55),
    illegal_injection = case_when(q56 == 1 ~ 0,
                                  q56 == 2 ~ 20,
                                  q56 == 3 ~ 40),
    drug_use_sum = rowSums(across(marijuana:illegal_injection)),
    other_drug_sum = rowSums(across(cocaine:illegal_injection)),
    drug_use_freq = drug_use_frequency(drug_use_sum),
    drunk_driving = ifelse(q9 > 1 | q10 > 2, "Yes", "No"),
    drunk_driving = factor(drunk_driving, level = c("Yes", "No")),
    text_driving = ifelse(q11 == 2, "No", "Yes"),
    text_driving = factor(text_driving, level = c("Yes", "No")),
    carrying_weapon = ifelse(q14 == 1, "No", "Yes"),
    carrying_weapon = factor(carrying_weapon, level = c("Yes", "No")),
    suicide_attempt = ifelse(q26 == 1, "Yes", "No"),
    suicide_attempt = factor(suicide_attempt, level = c("Yes", "No")),
    quit_smoke = case_when(q39 == 1 ~ "Never Smoke",
                           q39 == 2 ~ "Yes",
                           q39 == 3 ~ "No"),
    quit_smoke = factor(quit_smoke, level = c("Never Smoke", "Yes", "No")),
    physical_fight = ifelse(q17 == 1, "No", "Yes"),
    physical_fight = factor(physical_fight, level = c("Yes", "No")),
    early_sex = ifelse(q59 > 1 & q59 <= 5 , "Yes", "No"),
    early_sex = factor(early_sex, level = c("Yes", "No")),
    q79 = as.numeric(q79),
    q80 = as.numeric(q80),
    tv_use = case_when(q79 == 1 ~ 0, 
                       q79 == 2 ~ 0.5,
                       q79 > 2 ~ q79 - 2),
    computer_use = q80 - 2,
    computer_use = case_when(q80 == 1 ~ 0, 
                       q80 == 2 ~ 0.5,
                       q80 > 2 ~ q80 - 2),
    screening_use = tv_use + computer_use,
    sleeping_time = as.numeric(q88) + 3,
    smoking_status = case_when(q33 == 1 ~ "Never Smoker",
                               q33 >= 2 & q33 <= 5 ~ "Light Smoker",
                               q33 >=6 ~ "Heavy Smoker"),
    smoking_status = factor(smoking_status, level = c("Never Smoker", "Light Smoker", "Heavy Smoker")),
    binge_drinking = case_when(q42 == 1 ~ "No Binge Drinking",
                               q42 >= 2 & q42 <= 4 ~ "Light Binge Drinking",
                               q42 >=5 ~ "Heavy Binge Drinking"),
    binge_drinking = factor(binge_drinking, level = c("No Binge Drinking", "Light Binge Drinking", "Heavy Binge Drinking")),
    seat_belt = case_when(q8 <= 2 ~ "Never or Rarely",
                          q8 == 3 ~ "Sometimes",
                          q8 >= 4 ~ "Most of the time or Always"),
    seat_belt = factor(seat_belt, level = c("Never or Rarely", "Sometimes", "Most of the time or Always")),
    grades_school = case_when(q89 == 1 ~ "Mostly A's",
                       q89 == 2 ~ "Mostly B's",
                       q89 == 3 ~ "Mostly C's",
                       q89 %in% c(4,5) ~ "Mostly Below C's",
                       q89 == 6 ~ "None of these grades",
                       q89 == 7 ~ "Not sure"),
    grades_school = factor(grades_school, level = c("Mostly A's", "Mostly B's", "Mostly C's", "Mostly Below C's", "None of these grades", "Not sure")),
    smoke_age = firstage_tidy(q31),
    alcohol_age = firstage_tidy(q40)
    ) %>% 
  select(drug_use_freq, grade, sex, bmi, race4, drunk_driving, text_driving, carrying_weapon, suicide_attempt, quit_smoke, physical_fight, early_sex, screening_use, sleeping_time, smoking_status, binge_drinking, seat_belt, grades_school, smoke_age, alcohol_age) 
```

This data analysis uses the resulted data set cleaned by the processing described in part 2 of [Data]("3,Data Description.html"). According to our motivation, we are interested the relationship between drug use and youth behaviors. In this analysis, we will explore the association between drug use status and each health behavior variables.

There are three kinds of health-related variables and we will use different methods to analyze: 
 - Bi-level categorical variables: drunk driving, carrying weapon, suicide attempt, physical fight, quit-smoking, early-age sexual intercourse, heavy screening usage, enough sleeping
 - Multivariate categorical variables: smoking status, binge drinking status, seat belt usage, GPA
 - Quantitative Variables: initial smoking age, initial drinking age, BMI
 
Here is a table summarized the overall distribution of important variables by drug use status.

## Three-line-table

```{r three-line}

label(yrbss_state_tidy_additional$drunk_driving) = "Drunk Driving"
label(yrbss_state_tidy_additional$text_driving) = "Text Driving"
label(yrbss_state_tidy_additional$carrying_weapon) = "Weapon Carrying"
label(yrbss_state_tidy_additional$suicide_attempt) = "Suicide Attempt"
label(yrbss_state_tidy_additional$quit_smoke) = "Quit Smoking"
label(yrbss_state_tidy_additional$physical_fight) = "Physical Fight"
label(yrbss_state_tidy_additional$early_sex) = "Sexual Intercourse"
label(yrbss_state_tidy_additional$screening_use) = "Screening Use"
label(yrbss_state_tidy_additional$sleeping_time) = "Sleeping Time"
label(yrbss_state_tidy_additional$smoking_status) = "Smoking Status"
label(yrbss_state_tidy_additional$binge_drinking) = "Binge Dringking"
label(yrbss_state_tidy_additional$seat_belt) = "Seat Belt Use"
label(yrbss_state_tidy_additional$grades_school) = "Grades in School"


table1(~ drunk_driving + text_driving + carrying_weapon + suicide_attempt + quit_smoke + physical_fight + early_sex + screening_use + sleeping_time + smoking_status + binge_drinking + seat_belt + grades_school| drug_use_freq, data = yrbss_state_tidy, caption = "Table 1: Distribution of risky behaviors by drug use status")
```



## Hypothesis Testing with visualization of drug status comparison

### Chi-square Test

For bi-level and multi-level categorical variables, we test the independence of two categorical variables, so we should use the R X C chi-square test with hypotheses that 

$$
H_0:\text{There is no association between drug use and the health-related variable of interest}
H_1:\text{There is an association between drug use and the health-related variable of interest}
$$
Other than the chi-square test, we also use the Cramer's V Effect Size to measure the magnitude of the association. It is measure by the manner:
$$
V = \sqrt{\frac{X^2}{n*df}}
$$
where $X^2$ is the chi-square statistic, n is the sample size and $df$ is the degrees of freedom.\

Usually, we use the table below to assess the association through the Cramer's V effect size

```{r}
tibble(
  df = 1:5,
  small_effect = c(0.10, 0.07, 0.06, 0.05, 0.04),
  medium_effect = c(0.3, 0.21, 0.17, 0.15, 0.13),
  large_effect = c(0.5, 0.35, 0.29, 0.25, 0.22)
) %>% 
  knitr::kable()
```


Besides the crude analysis which test the association for all observations in the data set, we also perform a stratified analysis over grade, gender and race to investigate the interaction or confounding effect of these demographic variable on the crude association. 

### Drunk driving

```{r}
test_of_association(yrbss_state_tidy_additional, "drunk_driving", "Table 2: Drug Use Status vs Drunk Driving")
```


From Table 2, we can see that the overall p-value is significantly small and the p-value of each stratum is significantly small to say that there is association between drug use status and drunk driving. There are differences of effect size among grade, gender and race, implying some interaction effect, but the lowest effect size is still above 0.22, suggesting a moderate association between drug use status and the drunk driving. The effect size demonstrates that the association might be stronger among higher grade, male and white race.

Overall:

```{r}
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Drunk Driving",
       x = "Drug Use Status",
       title = "Plot 1: Proportion of Drunk Driving Among Drug Use Status")
```

Stratified Analysis:

```{r}
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(y = "Proportion of Drunk Driving",
       x = "Drug Use Status",
       fill = "Grade",
       title = "Plot 1.1: Proportion of Drunk Driving Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Drunk Driving",
       x = "Drug Use Status",
       fill = "Gender",
       title = "Plot 1.2: Proportion of Drunk Driving Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_drunk_driving = sum(drunk_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_drunk_driving, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Drunk Driving",
       x = "Drug Use Status",
       fill = "Race",
       title = "Plot 1.3: Proportion of Drunk Driving Among Drug Use Status")
```

All proportion plots display that an increasing level of drug use frequency has a higher proportion of drunk driving. The dose-response relationship also verifies the existence of the association.

### cell phone use while driving


```{r}
test_of_association(yrbss_state_tidy_additional, "text_driving", "Table 3: Drug Use Status vs Cell Phone Use During Driving")
```

From Table 3, we can see that not all p-values are significant enough to conclude an association between drug_use status and cell phone use while driving. Since younger children have much less chance of driving than older children, it is make sense that the association is heavily impacted by the grade. However, it also shows no association between drug_use status and cell phone use while driving among Black or African American and all other races. In a conservative perspective, we would conclude that the association between drug_use status and cell phone use while driving is weak, especially among Black or African American and all other races students.

```{r}
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Cell Phone Use During Driving",
       x = "Drug Use Status",
       title = "Plot 2: Proportion of Cell Phone Use During Driving Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Cell Phone Use During Driving",
       x = "Drug Use Status",
       title = "Plot 2.1: Proportion of Cell Phone Use During Driving Among Drug Use Status")

yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Cell Phone Use During Driving",
       x = "Drug Use Status",
       title = "Plot 2.2: Proportion of Cell Phone Use During Driving Among Drug Use Status")

yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_text_driving = sum(text_driving == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_text_driving, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Cell Phone Use During Driving",
       x = "Drug Use Status",
       title = "Plot 2.3: Proportion of Cell Phone Use During Driving Among Drug Use Status",
       fill = "race")
```

The overall proportion of cell phone use while driving rises while the level of drug use rises. However, the increment is small and invisible after stratifying by races. Female tends to have higher proportion of cell phone use while driving regardless of drug use status.


### Carrying weapon

```{r}
test_of_association(yrbss_state_tidy_additional, "carrying_weapon", "Table 4: Drug Use Status vs Carrying Weapon")
```

From Table 4, we can see that all p-values are significantly small to conclude that there is association between drug use status and carrying weapon. The overall effect size is around 0.26 while the lowest effect size is greater than 0.2, so we are confident to say that it is an medium association. However, the large difference of effect size among gender and race implies the interaction effects of gender and race on the association. The association between drug use status and carrying weapon is stronger among male compared to female; the association between drug use status and carrying weapon is strongest among Black or African American and weakest among White.

```{r}
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Carrying Weapon",
       x = "Drug Use Status",
       title = "Plot 3: Proportion of Carrying Weapon Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Carrying Weapon",
       x = "Drug Use Status",
       title = "Plot 3.1: Proportion of Carrying Weapon Among Drug Use Status")

yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Carrying Weapon",
       x = "Drug Use Status",
       title = "Plot 3.2: Proportion of Carrying Weapon Among Drug Use Status")

yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_carrying_weapon = sum(carrying_weapon == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_carrying_weapon, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Carrying Weapon",
       x = "Drug Use Status",
       title = "Plot 3.3: Proportion of Carrying Weapon Among Drug Use Status")
```

All plots demonstrate an rising trend of proportion of carrying weapon with the increase level of drug use frequency. The proportion of carrying weapon boosts from light dose drug use to heavy dose drug use. 9th Grade has highest proportion of carrying weapon regardless of drug use status while male has higher proportion of carrying weapon than female ar all drug use status. The proportion of Black and African American students carrying weapon suddenly escalates on heavy dose drug use.  


### Suicide Attempt


```{r}
test_of_association(yrbss_state_tidy_additional, "suicide_attempt", "Table 5: Drug Use Status vs Suicide Attempt")
```

From Table 5, we can see that all p-values are small enough, so we could reject the null hypothesis that there is association between drug use status and suicide attempt. The overall effect size is around 0.33 while the lowest effect size is greater than 0.21, so the association is moderately strong. The interaction effect of race in the association is detected by the difference of effect size among different races. The association is strongest for all other races while it is weakest for Black or African American.

```{r}
yrbss_state_tidy_additional %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Suicide Attempt",
       x = "Drug Use Status",
       title = "Plot 4: Proportion of Suicide Attempt Among Drug Use Status")

yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Suicide Attempt",
       x = "Drug Use Status",
       title = "Plot 4.1: Proportion of Suicide Attempt Among Drug Use Status")

yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Suicide Attempt",
       x = "Drug Use Status",
       title = "Plot 4.2: Proportion of Suicide Attempt Among Drug Use Status")

yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_suicide_attempt = sum(suicide_attempt == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_suicide_attempt, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Suicide Attempt",
       x = "Drug Use Status",
       title = "Plot 4.3: Proportion of Suicide Attempt Among Drug Use Status")
```

Based on these proportion plots, the proportion of suicide attempt increases while the level of drug use frequency goes up. This trend also shows up at each stratum in the stratified plot. The dose-response relationship supports our conclusion about the association between suicide attempt and drug use status. 


### Quit smoking


```{r}
yrbss_state_tidy_quit_smoke <- yrbss_state_tidy_additional %>% filter(quit_smoke != "Never Smoke")
test_of_association(yrbss_state_tidy_quit_smoke, "quit_smoke", "Table 6: Drug Use Status vs Quit Smoking")
```

For quit smoking, we should focus on students who smoke so the number of non-smokers won't bias our results. After filtering out the nonsmokers, the resulted table is in Table 6. Some p-values are not significantly small enough to conclude the association. The overall effect size is around 0.08 and the difference of effect size among grades and among races are significantly large, so the confounding and interaction effects heavily impacted the analysis of association. We would conclude that there is no association between drug use status and quit smoking.   


```{r}
yrbss_state_tidy_quit_smoke %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Quit Smoking",
       x = "Drug Use Status",
       title = "Plot 5: Proportion of Quit Smoking Among Drug Use Status")
  


yrbss_state_tidy_quit_smoke %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(y = "Proportion of Quit Smoking",
       x = "Drug Use Status",
       title = "Plot 5.1: Proportion of Quit Smoking Among Drug Use Status")

yrbss_state_tidy_quit_smoke %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Quit Smoking",
       x = "Drug Use Status",
       title = "Plot 5.2: Proportion of Quit Smoking Among Drug Use Status")

yrbss_state_tidy_quit_smoke %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_quit_smoke = sum(quit_smoke == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_quit_smoke, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Quit Smoking",
       x = "Drug Use Status",
       title = "Plot 5.3: Proportion of Quit Smoking Among Drug Use Status")
```

The proportion bar plots do not demonstrate any dose-response relationship since the proportion has no trend of increasing or decreasing across the drug use status. Moreover, the difference of proportion is small between each drug use status. Thus, these plots support the conclusion in the analysis table that there is no association between quit smoking and drug use status.


### physical fight



```{r}
test_of_association(yrbss_state_tidy_additional, "physical_fight", "Table 7: Drug Use Status vs Physical Fight")
```

From Table 7, we can see that all p-value are significantly small and the effect size are around 0.4, it is the association between drug use status and physical fight is strong. The effect size over each stratum is very close to each other, so the interaction effect is negligible. 

```{r}
yrbss_state_tidy_additional %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Physical Fight",
       x = "Drug Use Status",
       title = "Plot 6: Proportion of Physical Fight Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Physical Fight",
       x = "Drug Use Status",
       title = "Plot 6.1: Proportion of Physical Fight Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Physical Fight",
       x = "Drug Use Status",
       title = "Plot 6.2: Proportion of Physical Fight Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_physical_fight = sum(physical_fight == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_physical_fight, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Physical Fight",
       x = "Drug Use Status",
       title = "Plot 6.3: Proportion of Physical Fight Among Drug Use Status")
```

The plots illustrate a positive relationship between drug use status and proportion of physical fight, so the proportion of physical fight is higher when the drug use is heavier. This positive trend is stable in each stratum in the stratified plot. The plots are consistent with our conclusion based on the statistical analysis table and enhance the inference about the association between drug use status and physical fight.


### Early-age Sexual Intercourse



```{r}
test_of_association(yrbss_state_tidy_additional, "early_sex", "Table 8: Drug Use Status vs Early-Age Sexual Intercourse")

yrbss_state_tidy_early_sex %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Early-Age Sexual Intercourse",
       x = "Drug Use Status",
       title = "Plot 7: Proportion of Early-Age Sexual Intercourse Among Drug Use Status")


yrbss_state_tidy_early_sex %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Early-Age Sexual Intercourse",
       x = "Drug Use Status",
       title = "Plot 7.1: Proportion of Early-Age Sexual Intercourse Among Drug Use Status")


yrbss_state_tidy_early_sex %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Early-Age Sexual Intercourse",
       x = "Drug Use Status",
       title = "Plot 7.2: Proportion of Early-Age Sexual Intercourse Among Drug Use Status")


yrbss_state_tidy_early_sex %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_early_sex = sum(early_sex == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_early_sex, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Early-Age Sexual Intercourse",
       x = "Drug Use Status",
       title = "Plot 7.3: Proportion of Early-Age Sexual Intercourse Among Drug Use Status")

```



### Screening Use

```{r}
yrbss_state_tidy_screening_use <- 
  yrbss_state_tidy %>%
  drop_na(q79, q80) %>% 
  mutate(
    q79 = as.numeric(q79),
    q80 = as.numeric(q80),
    tv_use = case_when(q79 == 1 ~ 0, 
                       q79 == 2 ~ 0.5,
                       q79 > 2 ~ q79 - 2),
    computer_use = q80 - 2,
    computer_use = case_when(q80 == 1 ~ 0, 
                       q80 == 2 ~ 0.5,
                       q80 > 2 ~ q80 - 2),
    screening_use = tv_use + computer_use
  )
yrbss_state_tidy_screening_use %>% 
  group_by(drug_use_freq, screening_use) %>% 
  count() %>% 
  group_by(drug_use_freq) %>% 
  mutate(
    prop = n/sum(n)
  ) %>% 
  ggplot(aes(x = screening_use, y = prop)) +
  geom_bar(stat = "identity") +
  geom_smooth(se = FALSE) +
  facet_grid(.~drug_use_freq)
```


```{r}
test_of_association(yrbss_state_tidy_additional, "heavy_screen_use", "Table 9: Drug Use Status vs Heavy Screening Use")

yrbss_state_tidy_additional %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Heavy Screening Use",
       x = "Drug Use Status",
       title = "Plot 8: Proportion of Heavy Screening Use Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Heavy Screening Use",
       x = "Drug Use Status",
       title = "Plot 8.1: Proportion of Heavy Screening Use Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Heavy Screening Use",
       x = "Drug Use Status",
       title = "Plot 8.2: Proportion of Heavy Screening Use Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_heavy_screen_use = sum(heavy_screen_use == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_heavy_screen_use, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Heavy Screening Use",
       x = "Drug Use Status",
       title = "Plot 8.3: Proportion of Heavy Screening Use Among Drug Use Status")
```


### Sleeping time

```{r}
yrbss_state_tidy_sleeping_time <- 
  yrbss_state_tidy %>%
  drop_na(q88) %>% 
  mutate(
    sleeping_time = as.numeric(q88) + 3
  )

yrbss_state_tidy_sleeping_time %>% 
  group_by(drug_use_freq, sleeping_time) %>% 
  count() %>% 
  group_by(drug_use_freq) %>% 
  mutate(
    prop = n/sum(n)
  ) %>% 
  ggplot(aes(x = sleeping_time, y = prop)) +
  geom_bar(stat = "identity") +
  geom_smooth(se = FALSE) +
  facet_grid(.~drug_use_freq)
```


```{r}
test_of_association(yrbss_state_tidy_additional, "enough_sleep", "Table 10: Drug Use Status vs Enough Sleeping")

yrbss_state_tidy_additional %>%
  group_by(drug_use_freq) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Enough Sleeping",
       x = "Drug Use Status",
       title = "Plot 9: Proportion of Enough Sleeping Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  mutate(grade = ifelse(is.na(grade), "Missing", as.character(grade))) %>% 
  group_by(drug_use_freq, grade) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep, fill = grade)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Enough Sleeping",
       x = "Drug Use Status",
       title = "Plot 9.1: Proportion of Enough Sleeping Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, sex) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Enough Sleeping",
       x = "Drug Use Status",
       title = "Plot 9.2: Proportion of Enough Sleeping Among Drug Use Status")


yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4) %>% 
  summarize(
    prop_enough_sleep = sum(enough_sleep == "Yes")/n()
  ) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_enough_sleep, fill = race4)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(y = "Proportion of Enough Sleeping",
       x = "Drug Use Status",
       title = "Plot 9.3: Proportion of Enough Sleeping Among Drug Use Status")
```

### Smoking Status

Based on question 33 in the 2019 YOUTH RISK BEHAVIOR SURVEY, the researchers care about whether students smoked more than 10 cigarettes per day on the days they smoked during the 30 days before the survey. Therefore, we define students who did not smoke as "Never Smoker", 1 to 10 cigarettes per day as "Light Smoker", greater than 10 cigarettes per day as "Heavy Smokers".

```{r}
smoking_status_crude = 
  association_analysis(yrbss_state_tidy_additional, "smoking_status", "crude") %>% 
  mutate(type = "crude")

# chi-square of grade
smoking_status_grade =
  yrbss_state_tidy_additional %>% 
      group_by(across(all_of("grade")), drug_use_freq, across(all_of("smoking_status"))) %>%
      count() %>% 
      pivot_wider(values_from = n, names_from = starts_with("smoking_status"), names_prefix = "smoking_status") %>% 
  mutate(`smoking_statusHeavy Smoker` = ifelse(is.na(`smoking_statusHeavy Smoker`), 0 , `smoking_statusHeavy Smoker`)) %>% 
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      nest(data = starts_with("smoking_status")) %>% 
      mutate(
      chisq_tests = map(.x = data, ~chisq.test(x = .x)),
      chisq_result = map(chisq_tests, broom::tidy),
      totaln = map(data, colSums),
      totaln = map_dbl(.x = totaln, ~reduce(.x, sum))) %>% 
      unnest(chisq_result) %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter)) %>% 
      select(stratum = starts_with("grade"), 
             statistic, p.value, 
             df = parameter, 
             cramer_v_effect_size) %>% 
  mutate(type = "grade")

# chi-square of gender
smoking_status_gender = 
  association_analysis(yrbss_state_tidy_additional, "smoking_status", "sex") %>% 
  mutate(type = "sex")

# chi-square of race
smoking_status_race = 
  yrbss_state_tidy_additional %>% 
      group_by(across(all_of("race4")), drug_use_freq, across(all_of("smoking_status"))) %>%
      count() %>% 
      pivot_wider(values_from = n, names_from = starts_with("smoking_status"), names_prefix = "smoking_status") %>% 
  mutate(`smoking_statusHeavy Smoker` = ifelse(is.na(`smoking_statusHeavy Smoker`), 0 , `smoking_statusHeavy Smoker`)) %>% 
      ungroup() %>% 
      select(-drug_use_freq) %>% 
      nest(data = starts_with("smoking_status")) %>% 
      mutate(
      chisq_tests = map(.x = data, ~chisq.test(x = .x)),
      chisq_result = map(chisq_tests, broom::tidy),
      totaln = map(data, colSums),
      totaln = map_dbl(.x = totaln, ~reduce(.x, sum))) %>% 
      unnest(chisq_result) %>% 
      mutate(cramer_v_effect_size = sqrt(statistic / totaln*parameter)) %>% 
      select(stratum = starts_with("race4"), 
             statistic, p.value, 
             df = parameter, 
             cramer_v_effect_size) %>% 
  mutate(type = "race")

output = 
  bind_rows(smoking_status_crude, smoking_status_grade,smoking_status_gender, smoking_status_race) %>% 
  mutate(type = factor(type, levels = c("crude","grade", "sex", "race")))
  
kbl(output[, -6], caption = "Table **: Drug Use Status vs Binge Drinking") %>%
    kable_paper("hover", full_width = F) %>% 
    pack_rows(index = table(output$type))
```

```{r}
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, smoking_status) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c('#00CC00', '#FFA500', '#FF4500')) + 
  labs(title = "Proportion of Smoking Status Among Drug Use Status",
       x = "Drug Use Status",
       y = "Proportion",
       fill = "Smoking Status") 
```

```{r}
# Stratified Analysis over Grade
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq,grade, smoking_status) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c('#00CC00', '#FFA500', '#FF4500')) + 
  facet_grid(. ~ grade) +
  labs(title = "Stratified Analysis of Smoking Stauts over Grade",
       x = "Drug Use Frequency",
       y = "Proportion") 
```

```{r}
# Over Gender
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq,sex, smoking_status,) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  scale_fill_manual(values = c('#00CC00', '#FFA500', '#FF4500')) + 
  facet_grid(. ~ sex)
  labs(title = "Stratified Analysis of Smoking Stauts over Gender",
       x = "Drug Use Frequency",
       y = "Proportion") 
```




```{r}
# Over race
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4, smoking_status) %>% 
  summarize(n = n()) %>% 
  mutate(prop_smoking_status = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_smoking_status, fill = smoking_status)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ race4) + 
  labs(title = "Stratified Analysis of Smoking Stauts over Race",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


### Binge Drinking 

Based on question 42 in the 2019 YOUTH RISK BEHAVIOR SURVEY, we define students who did not have binge drinking as "No Binge Drinking", 1 to 5 days of binge drinking in the past 30 days as "Light Binge Drinking", greater than 5days as "Heavy Binge Drinking".

```{r}
test_of_association(yrbss_state_tidy_additional, "binge_drinking", "Table **: Drug Use Status vs Binge Drinking")
```

```{r}
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Proportion of Binge Drinking Among Drug Use Status",
       x = "Drug Use Frequency",
       y = "Proportion")
```


```{r}
# Stratified Analysis over Grade
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq,grade, binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ grade) + 
  labs(title = "Stratified Analysis of Binge Drinking over Grade",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


```{r}
# Over Gender
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq,sex,binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ sex) +
  labs(title = "Stratified Analysis of Binge Drinking over Gender",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


```{r}
# Over race
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4, binge_drinking) %>% 
  summarize(n = n()) %>% 
  mutate(prop_binge_drinking = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_binge_drinking, fill = binge_drinking)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ race4) + 
  labs(title = "Stratified Analysis of Binge Drinking over Race",
       x = "Drug Use Frequency",
       y = "Proportion") 
```


### Seat belt use 
Resource: question 8

```{r}
test_of_association(yrbss_state_tidy_additional, "seat_belt", "Table **: Drug Use Status vs Seat Belt Use")

```

```{r}
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") + 
  labs(title = "Proportion of Seat Belt Use Among Drug Use Frequency",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Stratified Analysis over Grade
yrbss_state_tidy_additional%>% 
  group_by(drug_use_freq,grade, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ grade) +
  labs(title = "Stratified Analysis of Seat Belt Use over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Over Gender
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq,sex, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ sex) +
  labs(title = "Stratified Analysis of Seat Belt Use over Gender",
       x = "Drug Use Frequency",
       y = "Proportion") 
```



```{r}
# Over race
yrbss_state_tidy_additional %>% 
  group_by(drug_use_freq, race4, seat_belt) %>% 
  summarize(n = n()) %>% 
  mutate(prop_seat_belt = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_seat_belt, fill = seat_belt)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ race4) + 
  labs(title = "Stratified Analysis of Seat Belt Use over Race",
       x = "Drug Use Frequency",
       y = "Proportion")
```

### Grades in school 
Resource: question 89
exclude option "None of these grades" and "Not sure"

```{r}
yrbss_state_tidy_additional_grades_school <- 
  yrbss_state_tidy_additional %>% 
  filter(grades_school %in% c("Mostly A's", "Mostly B's", "Mostly C's","Mostly Below C's"))
```

```{r}
test_of_association(yrbss_state_tidy_additional_grades_school, "grades_school", "Table **: Drug Use Status vs Grades in School")
```

```{r}
yrbss_state_tidy_additional_grades_school %>% 
  group_by(drug_use_freq, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Proportion of Grades in School Among Drug Use Frequency",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Stratified Analysis over Grade
yrbss_state_tidy_additional_grades_school %>% 
  group_by(drug_use_freq,grade, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(.~ grade) +
  labs(title = "Stratified Analysis of Grades in School over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Over Gender
yrbss_state_tidy_additional_grades_school %>% 
  group_by(drug_use_freq,sex, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid(. ~ sex) +
  labs(title = "Stratified Analysis of Grades in School over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```

```{r}
# Over race
yrbss_state_tidy_additional_grades_school %>% 
  group_by(drug_use_freq, race4, grades_school) %>% 
  summarize(n = n()) %>% 
  mutate(prop_grades_school = n/sum(n)) %>% 
  ggplot(aes(x = drug_use_freq, y = prop_grades_school, fill = grades_school)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_grid( . ~ race4) +
  labs(title = "Stratified Analysis of Grades in School over Grade",
       x = "Drug Use Frequency",
       y = "Proportion")
```




## Part 3
##### 2，额外图  
1，吸毒各个frequency段的，吸烟初始年龄分布。  
2, .....................，喝酒初始年龄分布。  
3, ....................., bmi box plot 分男女 or 金字塔图

## Q1
```{r}
#tidy q31 smoking and q40 alcohol start age
firstage_tidy <- function(input_list){
  output <- case_when(input_list == 1 ~ "never",
                      input_list == 2 ~ "8 years old or younger",
                      input_list == 3 ~ "9 or 10 years old",
                      input_list == 4 ~ "11 or 12 years old",
                      input_list == 5 ~ "13 or 14 years old",
                      input_list == 6 ~ "15 or 16 years old",
                      input_list == 7 ~ "17 years old or older")
  return(output)
}

#function for plotting first smoke_age
drug1_smoke=
  yrbss_state_tidy %>%
  mutate(
    smoke_age = firstage_tidy(q31),
  ) %>%
  drop_na(q31)


drug1_smoke %>%
  group_by(drug_use_freq, smoke_age) %>%
  filter(smoke_age != "never") %>%
  summarize(n = n()) %>% 
  mutate(percentage = n/sum(n),
         percent = scales::percent(percentage),
         smoke_age = factor(smoke_age, 
                            levels = c("never",
                                       "8 years old or younger",
                                       "9 or 10 years old",
                                       "11 or 12 years old",
                                       "13 or 14 years old",
                                       "15 or 16 years old",
                                       "17 years old or older"))) %>% 
  ggplot(
       aes(x = drug_use_freq, y = percentage,
           fill = smoke_age)
       ) +
  geom_bar(stat = "identity",                                     
           position = "dodge") +
  scale_y_continuous(breaks = seq(0, 2, .1), 
                     label = scales::percent) +
  geom_text(aes(label = percent),                                     
            size = 3,
            position = position_dodge(0.9)) +
  scale_fill_brewer(palette = "Set2")+
  labs(title="smoke start age")


#boxplot
ggplot(drug1_smoke, 
       aes(x = factor(drug_use_freq), 
           y = age_initial_smoking,fill=factor(drug_use_freq))) +     
  geom_boxplot() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1),
    legend.position = "none") +
  labs(title="smoke start age")
```

## Q2-alcohol
```{r}
#function for plotting first alcohol_age
drug1_alcohol=
  yrbss_state_tidy %>%
  mutate(
    alcohol_age = firstage_tidy(q40)
  ) %>%
  drop_na(q40) %>%
  drop_na(drug_use_freq)


drug1_alcohol %>%
  group_by(drug_use_freq, alcohol_age) %>%
    filter(alcohol_age != "never") %>%
  summarize(n = n()) %>% 
  mutate(percentage = n/sum(n),
         percent = scales::percent(percentage),
         alcohol_age = factor(alcohol_age, 
                            levels = c("never",
                                       "8 years old or younger",
                                       "9 or 10 years old",
                                       "11 or 12 years old",
                                       "13 or 14 years old",
                                       "15 or 16 years old",
                                       "17 years old or older"))) %>% 
  ggplot(
       aes(x = drug_use_freq, y = percentage,
           fill = alcohol_age)
       ) +
  geom_bar(stat = "identity",                                     
           position = "dodge") +
  scale_y_continuous(breaks = seq(0, 2, .1), 
                     label = scales::percent) +
  geom_text(aes(label = percent),                                     
            size = 3,
            position = position_dodge(0.9)) +
  scale_fill_brewer(palette = "Set2")+
  labs(title="alcohol start age")

#boxplot
ggplot(drug1_alcohol, 
       aes(x = factor(drug_use_freq), 
           y = age_initial_alcohol,fill=factor(drug_use_freq))) +        
  geom_boxplot() +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1),
    legend.position = "none") +
  labs(title="alcohol start age")

```

## Q3-BMI
```{r}
BMI =
  yrbss_state_tidy %>%
  drop_na(drug_use_freq) %>%
  drop_na(sex)

BMI %>%
  ggplot(aes(x = factor(drug_use_freq), 
           y = bmi,fill=factor(drug_use_freq))) +
  geom_boxplot() +
  facet_grid(. ~ sex) +  
  theme(
    axis.text.x = element_text(angle = 45, vjust = 0.5, hjust = 1),
    legend.position = "none")+
  labs(title = "Salary distribution by rank")

#heavy tail - more dose have higher BMI 


#pyramid plot - looks weird
ggplot(BMI, aes(x = drug_use_freq, y = ifelse(sex == "2", -bmi, bmi), fill =sex)) + 
  geom_bar(stat = "identity") +
  scale_y_continuous(labels = abs, limits = c(-200,200)) +
  coord_flip()

```



## Logistic regression
